\begin{thebibliography}{9}
\bibitem{zyk_pol}
\textit{Sergey Zykin, Andrey Poluyanov}
Multidimensional data building using Intermediate Representations.
"Administration problems".
2013.
NO. 5, P. 54--59.

\bibitem{chang}
\textit{Chang-Sup Park, Myoung Ho Kim b, Yoon-Joon Lee}
Usability-based caching of query results in OLAP systems.
// The Journal of Systems and Software
2003.
Vol. 68, - P. 103--119

\bibitem{baralis}
\textit{Baralis, E., Paraboschi, S., Teniente, E.}
Materialized view selection in a multidimensional database.
// Proc. of the 23rd International Conference on Very Large Data Bases,
Athens, Greece.
1997.
- P. 318--329

\bibitem{kalnis}
\textit{Kalnis, P., Papadias, D.}
Proxy-server architectures for OLAP.
// Proceedings of the 2001 ACM SIGMOD International Conference on Management of Data.
- Santa Barbara, CA, 2001.
- P. 367--378.

\bibitem{Afrati06}
\textit{Afrati F., Li C., Mitra P.}
Rewriting queries using views in the presence of arithmetic comparisons
// Theoretical Computer Science.
- 2006.
- Vol. 368.
- P. 88--123.

\bibitem{Keller96}
\textit{Keller M., Basu J.}
A predicate-based caching scheme for client-server database architectures
// VLDB Journal.
- 1996. N 5.
- P. 35--47.

\bibitem{shim}
\textit{Shim, J., Scheuermann, P., Vingralek R.}
Dynamic caching of query results for decision support systems.
// Proceedings of the 11th International Conference on Scientific and Statistical Database Management.
- Cleveland, OH, 1999.
- P. 254--263.

\end{thebibliography}